A comparison of classification algorithms applied to the Adult Database

### 2009-12-09 report/final.tex

>>>>

\subsection{Random Forest}

In Weka, the Random Forest algorithm can be parametrized by the number
of trees to build \textbf{$<$-I integer$>$}. Table \ref{tbl:results_random_forest} shows results for
a varying number of trees.

From the table we conclude that we need a certain amount of trees ($\approx$80) to
reach a satisfactory error rate ($<$22\%) for the test set. Interestingly
the error rate for the train set is low ($\approx$9\%), which indicates over-fitting.

One drawback of using a big number of trees is the use of large
quantities of memory (we needed 1GB to compute with 400 trees) and the time needed to compute everything.

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} \\ \hline
    -I 2 & 10.52 & 34.03 & 0m25.861s \\ \hline
    -I 3 & 9.83 & 34.00 & 0m26.318s \\ \hline
    -I 4 & 9.69 & 30.91 & 0m26.154s \\ \hline
    -I 5 & 9.50 & 29.62 & 0m27.108s \\ \hline
    -I 7 & 9.32 & 28.03 & 0m27.158s \\ \hline
    -I 10 & 9.24 & 29.80 & 0m28.107s \\ \hline
    -I 15 & 9.18 & 28.80 & 0m30.716s \\ \hline
    -I 20 & 9.15 & 28.09 & 0m36.340s \\ \hline
    -I 40 & 9.15 & 23.43 & 0m39.687s \\ \hline
    -I 60 & 9.15 & 23.65 & 0m49.443s \\ \hline
    -I 80 & 9.15 & 22.92 & 0m57.652s \\ \hline
    -I 100 & 9.15 & 22.83 & 1m28.024s \\ \hline
    -I 400 & 9.15 & 22.71 & 4m58.842s \\ \hline
  \end{tabular}
  \caption{Results for Random Forest.}
  \label{tbl:results_random_forest}
  \end{center}
\end{table}

\subsection{CART}

The simplified version of the CART algorithm is available
in the \textit{SimpleCart} Weka class.

This algorithm allows the disabling of various optimizations
and heuristics \textbf{$<$-U and -H $>$} as explained by the Weka manual.
We will keep these optimizations in.

Other interesting parameters that we will be changing, include:

\begin{itemize}
  \item \textbf{$<$-M int$>$}: to set the minimal number of instances at the terminal nodes.
  \item \textbf{$<$-N int$>$}: to set the number of folds used in the minimal cost-complexity pruning.
\end{itemize}

\begin{table}[ht]
  \begin{center}
  \begin{tabular}{ | l | c | c | c | c | c |}
    \hline
    \textbf{Parameters} & \textbf{Train error rate} & \textbf{Test error rate} & \textbf{Time} & \textbf{Nodes} & \textbf{Leaves} \\ \hline
    -M 2 -N 5 & 13.96 & 25.37 & 3m6.315s & 119 & 60 \\ \hline
    -M 2 -N 10 & 13.85 & 25.38 & 5m36.925s & 139 & 70 \\ \hline
    -M 2 -N 15 & 13.93 & 25.41 & 8m1.214s & 123 & 62 \\ \hline

  \end{tabular}
  \caption{Results for CART.}
  \label{tbl:results_cart}
  \end{center}
\end{table}

CART results are displayed in Table \ref{tbl:results_cart}.
